Formal Analysis of Rate-Monotonic Scheduling Implementation via Real-Time Maude

### 2015-10-10 RMS/submission.tex
first completed version

Liu and Layland also gave a sufficient condition for the
schedulability of a set of $n$ tasks scheduled by RMS algorithm
in~\cite{DBLP:journals/jacm/LiuL73}:
$\displaystyle\Sigma^n_{i=1}C_i/T_i \le n(2^{1/n}-1)$, where $C_i$ and
$T_i$ are the computation time and the period of task $\tau_i$
respectively. Two main directions on RMS have been explored since
then. One is to relax the assumptions on the original RMS model, for
instance, to allow aperiodic
tasks~\cite{DBLP:conf/rtss/LehoczkySS87,DBLP:journals/rts/SpruntSL89,DBLP:conf/rtss/LehoczkyR92,DBLP:journals/tc/StrosniderLS95},
to generalize it to be
\emph{deadline-monotonic}~\cite{DBLP:journals/pe/LeungW82,audsley1993deadline},
to allow resource sharing among tasks~\cite{DBLP:journals/tc/ShaRL90},
to extend it on
multiprocessors~\cite{dhall1978real,DBLP:journals/rts/LopezGDG03,DBLP:journals/tpds/LopezDG04,DBLP:journals/tc/BaruahG03},
to enhance
fault-tolerance~\cite{DBLP:journals/rts/OhS94,DBLP:journals/rts/GhoshMMS98,DBLP:journals/tpds/BertossiMR99}
and so on, making RMS algorithm applicable on more systems. The other
direction is to generate better schedulablity test conditions for the
algorithm and its
extensions~\cite{DBLP:conf/rtss/LehoczkySD89,DBLP:conf/rtss/KuoM91,DBLP:journals/tc/BiniBB03,DBLP:journals/rts/LopezGDG03,DBLP:journals/tc/BaruahG03,gardner1999}. The
RMS algorithm is of such practical importance that we certainly cannot
exhaust all the results in literatures here.

@@

\section{Related Work}
\label{s:relate}

\section{Conclusion}
\label{s:conclusion}

>>>>

Liu and Layland also gave a sufficient condition for the
schedulability of a set of $n$ tasks scheduled by RMS algorithm
in~\cite{DBLP:journals/jacm/LiuL73}:
$\displaystyle\Sigma^n_{i=1}C_i/T_i \le n(2^{1/n}-1)$, where $C_i$ and
$T_i$ are the computation time and the period of task $\tau_i$
respectively. Two main directions on RMS have been explored since
then. One is to relax the assumptions on the original RMS model, for
instance, to allow aperiodic
tasks~\cite{DBLP:conf/rtss/LehoczkySS87,DBLP:journals/rts/SpruntSL89,DBLP:conf/rtss/LehoczkyR92,DBLP:journals/tc/StrosniderLS95},
to generalize it to be
\emph{deadline-monotonic}~\cite{DBLP:journals/pe/LeungW82,audsley1993deadline},
to allow resource sharing among tasks~\cite{DBLP:journals/tc/ShaRL90},
to extend it on
multiprocessors~\cite{dhall1978real,DBLP:journals/rts/LopezGDG03,DBLP:journals/tpds/LopezDG04,DBLP:journals/tc/BaruahG03},
to enhance
fault-tolerance~\cite{DBLP:journals/rts/OhS94,DBLP:journals/rts/GhoshMMS98,DBLP:journals/tpds/BertossiMR99}
and so on, making RMS algorithm applicable on more systems. The other
direction is to generate better schedulablity test conditions for the
algorithm and its
extensions~\cite{DBLP:conf/rtss/LehoczkySD89,DBLP:conf/rtss/KuoM91,DBLP:journals/tc/BiniBB03,DBLP:journals/rts/LopezGDG03,DBLP:journals/tc/BaruahG03,gardner1999}. The
RMS algorithm is of such practical importance that we certainly cannot
exhaust all the results in the literature here.

@@

\section{Related Work}
\label{s:relate}
In this section we discuss our results with related work in two
directions.

Considering schedulability test, Liu and Layland gave
in~\cite{DBLP:journals/jacm/LiuL73} the famous sufficient condition
that a set of periodic tasks is schedulable with respect to RMS if
$\displaystyle \Sigma^n_{i=1} C_i/T_i \le n(2^{1/n}-1)$ holds, where
$C_i$ and $T_i$ are the computation time and the period of task
$\tau_i$ respectively. Then in~\cite{DBLP:journals/tc/BiniBB03}, a
more sufficient condition, known as \emph{Hyperbolic Bound}, which has
the same complexity as Liu and Layland's one, was proposed by Bini et
al. On the other hand, necessary and sufficient conditions for
schedulability were derived independently
in~\cite{DBLP:journals/rts/SpruntSL89} and~\cite{audsley1993deadline},
requiring more sophisticated analysis on the task set. Nevertheless,
these results take no overhead into account, being not as realistic as
ours. Katcher et al. did consider overhead in their schedulability
analysis under several models based on different kinds of popular
implementations~\cite{DBLP:journals/tse/KatcherAS93}. But as we
mentioned at the very beginning, the ways to implement the RMS
algorithm are countless, hence their analysis could not be applied on
our target implementation. Furthermore, comparing with those
theoretical analysis, our method by formal modeling and verification
has a great advantage that if our schedulability test answers ``no'',
it returns at the same time a real counterexample, which is able to
guide our engineer to adjust the design.

\cite{DBLP:conf/iceccs/CuiDT14} and~\cite{TianD2011} also made use of
model checking to analyze the RMS algorithm along the same line with
different languages and tools. They only considered the ideal setting
as in~\cite{DBLP:journals/jacm/LiuL73}, but not an implementation
which contains much more details. Our model of the implementation
easily degenerates to a model of the RMS algorithm, if we let the time
for scheduling and switching be zero and let the possible finishing
interval of each task be a single value.

Finally, Maude and Real-Time Maude have been successfully applied on
large numbers of applications~\cite{DBLP:journals/jlp/Meseguer12},
especially on communication protocols, real-time and cyber-physical
systems. But few results are achieved on scheduling problems. RMS is
investigated using Real-Time Maude for the first time.

\section{Conclusion}
\label{s:conclusion}
We have formally modeled a realistic implementation of RMS algorithm,
obtained from our industrial partner, using Real-Time Maude, a
modeling language based on rewriting logic. By taking into acount the
overhead of scheduling and switching, assuming the possible
computation time for each task may not be constant, we believe that
our model has enough details to be analyzed for the behaviours of the
real system. Two important properties--schedulability and
correctness--are verified by model checking on the model, under
several given scenarios. We also prove at last the soundness and
completeness of our analysis.
